Long term cloud enhancement events of global solar irradiance over Thessaloniki

### 2024-05-26 Article_B5/Article_B5.Rmd
Commit 2024-05-27 00:14

author:
  - name:        Athanasios N. Natsis
    email:       natsisphysicist@gmail.com
    affiliation: LAP
    correspondingauthor: true
    footnote: 1
  - name:        Alkiviadis Bais
    email:       abais@auth.gr
    affiliation: LAP
  - name:        Charikleia Meleti
    email:       meleti@auth.gr
    affiliation: LAP
    footnote: 2
address:
  - code:         LAP
    organization: Laboratory of Atmospheric Physics, Physics Department, Aristotle University of Thessaloniki
    # addressline:  1 main street
    city:         Thessaloniki
    postcode:     54124
    country:      Greece
  - code:         Another University
    organization: Department
    addressline:  A street 29
    postcode:     2054 NX
    city:         Manchester,
    state:        State
    country:      The Netherlands
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Here wii be the abstract.

  It consists of two paragraphs.

@@

keywords: 
  - cloud enhancement
  - total solar radiation
  - global horizonta irradiance
  - over irradiance

@@

Some of the proposed underling mechanisms of those events, have been summarized by
@Gueymard2017, and include scattering on the edge of cumulus clouds or through thin
cirrus.  Further investigation with radiation transfer modeling methods and
observations have pointed as the prevailing mechanism, the forward Mie scattering
[@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015], through the clouds.
The overall phenomenon depends on different interactive factors, that include cloud
thickness, constitution and type; and the relative position of the sun, the clouds,
and the observer [@Gueymard2017; @Veerman2022].  As such, there are multiple
contributing mechanism that are responsible for the observed irradiance enhancements.

Cloud enhancements have been reported, to be able to exceed in intensity even the
value of the solar constant, resulting of clear indexes above unit. A summary of
extreme enchantment cases have been compiled by @Martins2022. There are also some
practical implication of the cloud enchantments.  The intensity and duration of
enhancements can effect the efficiency and stability of photovoltaic power production
[@Lappalainen2020; @Jaervelae2020], and extreme enhancements cases, have the
potential to compromise the integrity of photovoltaic plants infrastructure
[@DoNascimento2019].  It has also been demonstrated, that these events can interfere
in the comparison of ground data and satellite observations [@Damiani2018]

<!--
Study of the mechanism has also been done with modeling of different cloud types and
atmospheric conditions.  @Gueymard2017; @Veerman2022
-->

<!--
Instruments respose ....
@Martins2022
-->

@@

<!-- our work -->
In this study, we evaluated the effects cloud enhancements on the total downward
radiation by studding the occurrences, their intensity, and their duration in a
thirty-year period at the city of Thessaloniki.  We used modeled clear sky
irradiance, as a baseline to identify cloud enhancements.  We were able to determine
some trends of the phenomenon, it's Climatology, and some of their  general
characteristics. We weren't able to find a comparable study that provides trends for
similar long term dataset, as ours.  The recording of the radiation signal is the
mean of one-minute. Thus, the minimum resolution of cloud enhancement events (CE), in
this study is one minute.

<!-- Define CE and ECE -->
In the relative bibliography different definition have been used for these events,
some of the are summarized by @Gueymard2017.  Here, we defined as cloud enhancement
events (CE) the cases when the measured global horizontal irradiance (GHI) at ground
level, exceeds the expected value under clear-sky conditions.  Similar, we define as
extreme cloud enhancement events (ECE), the cases when GHI, exceeds the Total Solar
Irradiance on horizontal plane at ground level.  Although the duration of these
bursts varies, from instantaneous to several minutes, here we are constrained by the
recorded data, to one-minute steps.

@@

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the continuous operation of the instruments.
The manual screening was aided by a radiation data quality assurance procedure,
adjusted for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must
exist at least 45 minutes of valid measurements, including nighttime near sunrise and
sunset. Days with less than 5 valid hours are rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty when the Sun is near the
horizon, and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than 
$`r 90 - BIO_ELEVA`^\circ$.

@@

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, here is defined as the irradiance difference of the
measured one-minute $\text{GHI}_i$ from the CE identification criterion in
Equation\nobreakspace\ref{eq:CE4}
($\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}$).
To have an estimation, and a first insight on the phenomenon, we experimented with two
simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which is
as simple clear sky model, and we had already adjusted and had good fit with our data
[@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both cases
by using an appropriate relative threshold and/or an additional constant offset.  The
initial results, showed that we can detect a big portion of the CE events. These
results were helpful, and helped to establish some criteria to further improve the CE
identification. It was evident, by inspecting the daily plot of irradiance, that
changes on the atmospheric conditions introduced numerous false positive and false
negative results.  To produce a more accurate reference, we had to take into account
more factors that effect the clear sky radiation. So we used a radiation transfer
model.

@@

We approximated the expected clear sky GHI by using a radiation transfer model.  The
simulations was performed by the well established model Libradtran [@Emde2016], a
similar approach, was also used by @Vamvakas2020 for creating a clear sky reference.
Because of the lack of observational data for the whole period, we used some long
term climatological data for the main factors responsible for the attenuation of the
broadband downward solar radiation in the atmosphere. Which are mainly, the aerosols
and the water vapors.  Fortunately, our site participates in the Aerosol Robotic
Network (AERONET) [@Giles2019; @Buis1998], as we operate a Cimel photometer since
2003, collocated with the CM-21 pyranometer.  The mean monthly aerosol optical depth
(AOD) on different wavelengths is provided by AERONET, along with the equivalent
water column height in the atmosphere.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated for combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($wc$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

<!-- Lookup table -->
To create a look-up table, that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability, in our one-minute GHI measurements, we adapted
each modeled value by scaling the model's input spectrum integral, to the
corresponding total solar irradiance (TSI) provided by NOAA [@Coddington2005].  Also,
we account for the effect of the Earth -- Sun distance on the irradiance, by using
the distance calculated by the Astropy [@AstropyCollaboration2022] software library.
As needed, we interpolate the resulting irradiances to the exact SZA of our
measurements.  For each period of the year, we used the appropriate atmospheric
profile (afglms or afglmw).  Finally, we calculated the clear sky irradiance value at
the horizontal plane.  Thus, we were able to emulate different atmospheric condition
and levels of atmospheric clearness for the climatological conditions of the site.
With this method, the modeled clear sky irradiances can be directly compared to each
measured one-minute value of GHI, for different conditions of atmospheric clearness.

@@

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better describe this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available AERONET data.

To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both, we
have used the AOD at $500\,\text{nm}$, which was inferred by the available Ångström
coefficients, and we choose the SZA of $55^\circ$ as a representative value for all
runs.

According to @Kazadzis2007, in the period 1997 -- 2005, the mean AOD of
$340\,\text{nm}$ was calculated to $0.403$ with the use of a Brewer
spectrophotometer, with a linear trend of $-3.8±0.93\,\%$ per year.  This corresponds
to a AOD $340\,\text{nm}$ change of $0.0153$ per year.  Using the alpha Ångström
coefficient of $\alpha = 1.6$, this translates to a change in Ångström beta
coefficient of $\beta=0.00272$ per year (or $\beta=0.084$ in 1997 and $\beta=0.059$
in 2005).  By interpolating the resulting clear sky irradiance, we have a trend of
$+0.21\,\%$ per year for the clear sky irradiance.  For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation scheme. We
computed the trend of clear sky radiation for each month of the year, which gave us a
mean clear sky radiation trend of $+0.14\,\%$ per year. We applied this long term
change to the clear sky irradiance, in order to create a more realistic
representation of the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change of clear sky irradiance relevant to its climatological values."}
knitr::include_graphics("../images/P_CS_change-1.png")
```

@@

Our main focus was to positively identify over irradiance events related to CEs.  For
this reason we had to select an appropriate set of atmospheric parameterization, in
Libradtran, and define an relevant threshold for the CE identification.  The
selection of the atmospheric parameterization will provide us with a reference, that
has a good agreement with the clear sky conditions. Additional, with the
implementation of the long term change, as described above, it will be able to align
well with most of the natural variation of the physical quantity.  The selection of
adequate threshold, will also allow us to have a robust method to distinguish the CE
cases, which will compensate for the limited accuracy of the input data, and the
natural variability of the atmosphere.

@@

In the process of defining the above details, we implemented an empirical method, by
manual inspection of the CE identification, on specially selected days from the whole
dataset.  We used seven sets of selected days, with characteristics relevant to the
efficiency of the identification threshold.  These sets were random groups of about
20 to 30 days with the following
characteristics:
(a) the largest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) without clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) without clouds and without EC events,
(e) with at least $60\,\%$ of the day length without clouds and some EC events,
(h) random days and
(i) some manual selected days that were included during the manual inspection.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

@@

The definition of the CE events with this method, has a level of subjectivity. The
actual clear sky irradiance, is not known, and can only be approximated. Although,
because our main focus is to be able to identify the OIR created by the clouds, we
expect that we were able to detect all the major events of OIR due to CE.  Where,
minor enhancements may be not detected, are less significant for most of our results.

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE, where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and satisfy the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, and $i$ each of the one-minute observation.

To give a better understanding of the CE identification procedure, we provide an
example of the CE identification for a selected day in the Appendix
(Figure\nobreakspace{}\ref{fig:example-day}), where the use of the reference and the
threshold are shown, along with the relevant measurements.  Similar we provide an
example plot of one year data, where the relation of the GHI and clear sky reference
is shown, along with the OIR intensity also in the Appendix
Figure\nobreakspace{}\ref{fig:example-year}.

@@

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are ECE events.
The highest GHI recorder was 
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`.
The absolute stronger CE event had an OIR of 
$`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
`r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`.
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

## Trends

@@

## CE Climatology

Another interesting aspect of the CE cases occurrences, is their seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurancies}, we have a monthly box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs on June.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
For the rest months the occurrences gradually ramp between the maximum and minimum.
These seasonality is a combinations of the clouds occurrences, and the relevant
position of the sun during the year. Unfortunately, the lack of detailed data of cloud
formation, type and location is not allowing further analysis.

@@

The distribution of the CE over irradiance intensity follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an inverse relation
between the events frequency and events intensity. This is expected as, the stronger
the CE events become, the rarer the particular atmospheric and sun conditions will
occur. This fact is indicative to the magnitude and probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR is 
higher due to the location of the the city of Patras, $2.5^\circ$ closer to the equator.

@@

## Groups stats

In order to further study the characteristics of the CE events, we grouped the single
minute CE events to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can represent cloud enhancement conditions spanning a
duration of multiple minutes.  We have identified `r ST_G0[, .N]` groups, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude, that
although we detected some CEG with durations longer than an hour, about $80\,\%$ of
the CEG have a duration of less than 5 minutes.

@@

The relation between the duration and the mean OIR intensity of the groups have been
studied (Figure\nobreakspace{}\ref{fig:group-2d}). We observe that the GCE events
tend to have either long duration, or large intensity, events with strong enhancement
and long duration are very rare. A similar result on this relation, have been also
reported by @Zhang2018, on a study using a far higher sampling rate than ours.

```{r group-2d, echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

## Extreme CE above TSI

A characteristic of the CE events that is commonly reported and has some significance
on the solar energy production infrastructure are the extreme CE events (ECE). Where
solar irradiance exceeds the expected irradiance on top of the atmosphere, we have
defined this as ECE (Equation\nobreakspace{}\ref{eq:ECE}).

<!-- -  max value ECE -->
<!--   relative to other works max values -->

Analogous to Figure\nobreakspace{}\ref{fig:relative-month-occurancies} we have
computed the distribution of the number of occurrences of ECE events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. The most active period
for ECE events is in the spring and the start of the summer (March -- June), followed
by a period in the late fall (September and October).  This is probably an effect of
the weather characteristics of those periods where, there are continuous alternations
between clear sky periods and clouds.

>>>>

author:
  - name:        Athanasios N. Natsis
    email:       natsisphysicist@gmail.com
    affiliation: LAP
    correspondingauthor: true
    footnote: 1
  - name:        Alkiviadis Bais
    email:       abais@auth.gr
    affiliation: LAP
  - name:        Charikleia Meleti
    email:       meleti@auth.gr
    affiliation: LAP
    footnote: 1
address:
  - code:         LAP
    organization: Laboratory of Atmospheric Physics, Physics Department, Aristotle University of Thessaloniki
    # addressline:  1 main street
    city:         Thessaloniki
    postcode:     54124
    country:      Greece
  # - code:         Another University
  #   organization: Department
  #   addressline:  A street 29
  #   postcode:     2054 NX
  #   city:         Manchester,
  #   state:        State
  #   country:      The Netherlands
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Here will be the abstract.

@@

keywords: 
  - cloud enhancement
  - total solar radiation
  - global horizontal irradiance
  - over irradiance

@@

Some of the proposed underling mechanisms of those enhancements, have been summarized
by @Gueymard2017, and include scattering on the edge of cumulus clouds or through
thin cirrus.  Further investigation with radiation transfer modeling methods and
observations have pointed as the prevailing mechanism, the forward Mie scattering
[@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015], through the clouds.
The overall phenomenon depends on different interactive factors, which includes cloud
thickness, constitution and type; and the relative position of the sun and the clouds
[@Gueymard2017; @Veerman2022].  As such, there are multiple contributing mechanisms
that are responsible for the observed irradiance enhancements.

On multiple sites, cloud enhancements have been reported, to be able to briefly exceed the
value of the solar constant, resulting of clear indexes above unit. A summary of
extreme enchantment cases (ECE) have been compiled by @Martins2022. There are also, some
practical implication of the cloud enchantments.  The intensity and duration of
enhancements can effect the efficiency and stability of photovoltaic power production
[@Lappalainen2020; @Jaervelae2020], and extreme enhancements cases, have the
potential to compromise the integrity of photovoltaic plants infrastructure
[@DoNascimento2019].  It has also been demonstrated, that these events can interfere
in the comparison of ground data and satellite observations [@Damiani2018]

<!--
Study of the mechanism has also been done with modeling of different cloud types and
atmospheric conditions.  @Gueymard2017; @Veerman2022
-->
<!--
Instruments respose ....
@Martins2022
-->

@@

<!-- our work -->
In this study, we evaluated the effects CE on the GHI by studding the occurrences, their intensity, and their duration in a
thirty-year period at the city of Thessaloniki.  We used modeled clear sky
irradiance, as a baseline to identify cloud enhancements.  We were able to determine
some trends of the phenomenon, it's Climatology, and some of their  general
characteristics. We weren't able to find a comparable study that provides trends for
similar long term dataset, as ours.  The recording of the GHI signal is the
mean of one-minute. Thus, the minimum resolution of CE events, in
this study is one minute.

<!-- Define CE and ECE -->
In the relative bibliography different definition have been used for these events,
some of the are summarized by @Gueymard2017.  Here, we defined as CE
events, the cases when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions.  Similar, we define as
extreme cloud enhancement events (ECE), the cases when GHI, exceeds the Total Solar
Irradiance (TSI) on horizontal plane at ground level.  Although the duration of these
bursts varies, from instantaneous to several minutes, here we are constrained by the
recorded data, to one-minute steps.

@@

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed, in order to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the long operation of the instrument.
The manual screening was aided by a radiation data quality assurance procedure,
adjusted for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must
exist at least 45 minutes of valid measurements, including nighttime, when near sunrise and
sunset. Days with less than 5 valid hours were rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty, when the Sun is near the
horizon, and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than 
$`r 90 - BIO_ELEVA`^\circ$.

@@

To be able to detect the CE cases, we had to establish a baseline, above which we 
characterized each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, is defined as the irradiance difference of the
measured one-minute $\text{GHI}_i$ from the CE identification criterion
($\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}.
To have an estimation, and a first insight on the phenomenon, we experimented with
two simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which
is as simple clear sky model, and we had already adjusted and had good fit with our
data [@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both
cases by using an appropriate relative threshold and/or an additional constant
offset.  The initial results, showed that we can detect a big portion of the CE
events. These results were helpful, as they are independent from unknown factors, and
helped us to establish some criteria to further improve the CE identification. It was
evident, by inspecting the daily plot of irradiance, that changes on the atmospheric
conditions introduced numerous false positive and false negative results.  To produce
a more accurate reference, we had to take into account more factors that effect the
clear sky radiation. So we used a radiation transfer
model in order to include the effects of ADO and the water vapors.

@@

We approximated the expected clear sky GHI by using the Libradtran radiation transfer
model [@Emde2016], a similar approach, was also used by @Vamvakas2020 for creating a
clear sky reference for studding cloud enhancement events.  Main factors responsible
for the attenuation of the broadband downward solar radiation in the atmosphere are
the aerosols and the water vapors.  Because of the lack of observational data for the
whole period, we used some long term climatological data to recreate the whole
period.  Fortunately, our site participates in the Aerosol Robotic Network (AERONET)
[@Giles2019; @Buis1998], as there is in operation a Cimel photometer since 2003,
collocated with the CM-21 pyranometer.  The mean monthly AOD on different wavelengths
are provided by AERONET, along with the equivalent water column height in the
atmosphere.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

<!-- Lookup table -->
To create a look-up table, that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability, in our one-minute GHI measurements, we adapted
each modeled value by scaling the model's input spectrum integral, to the
corresponding TSI, provided by NOAA [@Coddington2005].  Also,
we account for the effect of the Earth -- Sun distance on the irradiance, by using
the distance calculated by the Astropy [@AstropyCollaboration2022] software library.
As needed, we interpolate the resulting irradiances to the SZA of our
measurements.  For each period of the year, we used the appropriate atmospheric
profile (afglms or afglmw).  Finally, we calculated the clear sky irradiance value at
the horizontal plane.  Thus, we were able to emulate different atmospheric condition
and levels of atmospheric clearness for the climatological conditions of the site.
With this method, the modeled clear sky irradiances can be directly compared to each
measured one-minute value of GHI, for different conditions of atmospheric clearness.

@@

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents, mainly AOD. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.

To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both
inputs, we
used the AOD at $500\,\text{nm}$, which was inferred by the available Ångström
coefficients. We choose the SZA of $55^\circ$ as a representative value for all the
runs.

According to @Kazadzis2007, in the period 1997 -- 2005, the mean AOD of
$340\,\text{nm}$ was calculated to $0.403$ with the use of a Brewer
spectrophotometer, with a linear trend of $-3.8±0.93\,\%$ per year.  This corresponds
to a change of $0.0153$ per year for the AOD of $340\,\text{nm}$ .  Using the alpha Ångström
coefficient of $\alpha = 1.6$, this translates to a change in Ångström beta
coefficient of $\beta=0.00272$ per year (or $\beta=0.084$ in 1997 and $\beta=0.059$
in 2005).  By interpolating the resulting clear sky irradiance, we have a trend of
$+0.21\,\%$ per year for the clear sky irradiance.  For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation scheme. We
computed the trend of clear sky radiation for each month of the year, which gave us a
mean clear sky radiation trend of $+0.14\,\%$ per year. We applied this long term
change to the clear sky irradiance, in order to create a more realistic
representation of the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change of the clear sky irradiance relevant to its climatological values, due to changed of AOD."}
knitr::include_graphics("../images/P_CS_change-1.png")
```

@@

Our main focus was to positively identify over irradiance events related to CEs.  For
this reason we had to select an appropriate set of atmospheric parameterization, in
Libradtran, and define an relevant threshold for the CE identification.  The
selection of the atmospheric parameterization will provide us with a reference, that
has a good agreement with the clear sky conditions. Additional, with the
implementation of the long term change of AOD, as described above, it will be able to align
well with most of the natural variation of the physical quantity.  The selection of
adequate threshold, will allow us to have a robust method to distinguish the CE
cases, which will compensate for the limited accuracy of the input data, and the
natural variability of the atmosphere.

@@

In the process of defining the above details, we implemented an empirical method, by
manual inspection of the CE identification results, on selected days from the whole
dataset.  We used seven sets of days, with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 to 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) random selected days and
(i) some manual selected days that were included during the manual inspection process.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

@@

The definition of the CE events with this method, has a level of subjectivity. The
actual clear sky irradiance, is not known, and can only be approximated. Although,
because our main focus is to be able to identify the OIR created by the clouds, we
expect that we were able to detect all the major events of OIR due to CE.  Where,
some CE events, with very low OIR, may be not detected, these are few and their magnitude
will not effect our results.

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE, where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, and $i$ each of the one-minute observation.

To give a better understanding of the CE identification procedure, we provided an
example of the CE identification for a selected day, in the Appendix
(Figure\nobreakspace{}\ref{fig:example-day}), where the use of the clear sky
reference and the identification threshold are shown, along with the relevant
measurements.  Similar, we provided an example plot of data for one year, where the
relation of the GHI and clear sky reference is shown, along with the OIR intensity
(Appendix Figure\nobreakspace{}\ref{fig:example-year}).

@@

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI measurements, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`, and had
the maximum value of OIR at $`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 1)`\,W/m^2$.
<!-- The absolute strongest CE event had an OIR of 
<!-- $`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
<!-- `r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`. -->
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

## Long term trends

@@

## Cloud enhancements climatology

Another interesting aspect of the CE cases occurrences, is their seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurancies}, we have the monthly box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs on June.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
For the rest months, the occurrences gradually ramp between the maximum and minimum.
This seasonality is a combinations of the clouds occurrences, and the relevant
position of the sun during each season of the year. Unfortunately, the lack of detailed data on cloud
formation, type and location is not allowing further analysis.

@@

The distribution of the OIR intensity, follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an inverse relation
between the CE events frequency and their intensity. This is expected, as the stronger
the CE events become, the rarer the particular atmospheric and sun conditions can
occur. This fact, is indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR were
higher due to the location of the city of Patras, $2.5^\circ$ closer to the equator.

@@

## Groups of cloud enhancement

In order to study the total duration of the CE events, we grouped the one
minute CE events to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can represent cloud enhancement conditions spanning a
duration of multiple minutes.  We have identified `r ST_G0[, .N]` groups of CE in the
whole period of study, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude, that
although we detected some CEG with durations longer than an hour, about $80\,\%$ of
the CEG have a duration of less than 5 minutes.

@@

The relation between the duration and the mean OIR intensity of the groups have also been
studied (Figure\nobreakspace{}\ref{fig:group-2d}). We observed that the GCE events
tend to have either long duration, or large intensity. Events with strong enhancement
and long duration are very rare. Similar results on this relation, have been
reported by @Zhang2018, on a study using a far higher sampling rate than ours.

```{r group-2d, echo=FALSE, fig.cap="Relation of the mean over irradiance and duration of GCE, where the color scale donotes the frequency of the respected events."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

## Extreme cloud enhancement events

An aspect of the CE events that is commonly reported and has some significance
on the solar energy production infrastructure are the extreme CE events (ECE). Where
solar irradiance exceeds the expected irradiance on top of the atmosphere
(Equation\nobreakspace{}\ref{eq:ECE}).
Analogous to Figure\nobreakspace{}\ref{fig:relative-month-occurancies} we have
computed the distribution of the number of occurrences of ECE events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. The most active period
for ECE events is in the spring and the start of the summer (March -- June), followed
by a period in the late fall (September and October).  This is probably related to 
the weather characteristics of these periods, where there are continuous alternations
between clear sky periods and clouds.
